Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-01-20 paper/main.tex

The key idea behind UMM is that the probabilistic Mallows model (MM) used in
UMM does not simply represent the current sample $S$. Instead, the model
assigns higher probability values to the permutations with known (or expected)
minimal fitness values. This is achieved by means of the uBorda ranking
(Section~\ref{sec:uborda}), which can be seen as replicating each permutation
$\sigma \in S$ a number of times inversely proportional to its fitness
$f(\sigma)$, i.e., it behaves as if there were many copies of the best
solutions and few of the bad ones. A property of UMM is that the
computational complexity of working with the original sample (Borda) and with the
transformed sample (uBorda) is the same, i.e., polynomial. The most important steps are the update
(\emph{learning}) of the parameters $\sigma_0$, $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}) and
the sampling from $MM(\sigma_0,\theta)$ (lines~\ref{line:sample}--\ref{line:distance}). % in Algorithm~\ref{alg:umm}).

>>>>

The key idea behind UMM is that the probabilistic Mallows model (MM) used in
UMM does not simply represent the current sample $S$. Instead, the model
assigns higher probability values to the permutations with known (or expected)
minimal fitness values. This is achieved by means of the uBorda ranking
(Section~\ref{sec:uborda}) with $w(\sigma) = \rho^{f(\sigma)}$ and $\rho \in (0,1)$, which %  can be seen as replicating each permutation
% $\sigma \in S$ a number of times inversely proportional to its fitness
% $f(\sigma)$, i.e., it 
behaves as if there were many copies of the best
solutions and few of the bad ones. A property of UMM is that the
computational complexity of working with the original sample (Borda) and with the
transformed sample (uBorda) is the same, i.e., polynomial. The most important steps are the update
(\emph{learning}) of the parameters $\sigma_0$, $\theta$ and $\rho$ (line~\ref{line:uborda} in Algorithm~\ref{alg:umm}) and
the sampling from $MM(\sigma_0,\theta)$ (lines~\ref{line:sample}--\ref{line:distance}). % in Algorithm~\ref{alg:umm}).
